### 2016-01-13 mod1/manuscript/table_mbb.tex
add in necessary changes before moving table note to a column

\begin{deluxetable}{lcc}[tbpH]
\tabletypesize{\scriptsize}
\tablecolumns{3}
\tablecaption{SED fitting results}
\tablehead{
\colhead{Parameters}      &
\colhead{Optically Thick} &
\colhead{Optically Thin}
}
\startdata
$\chi^2$                      & 2.25                     & 5.31                    \\
D.O.F                         & 2                        & 3                       \\
$T$                           & 63.1$^{+1.1}_{-1.3}$     & 52.0$^{+1.3}_{-1.2}$    \\
$\beta$                       & 1.9$^{+0.6}_{-0.5}$      & 0.7$^{+0.2}_{-0.3}$     \\
$\alpha$                      & 2.9$^{+0.3}_{-0.4}$      & 2.8$^{+0.2}_{-0.2}$     \\
$\lambda_0$\tna               & 248.7$^{+86.0}_{-123.8}$ & ---                     \\
$\lambda_{\rm peak}$ \tnb     & 254.7$^{+6.2}_{-6.1}$    & 301.4$^{+29.0}_{-30.1}$ \\
$f_{\rm norm,\ 500\mu m}$\tnc & 267.4$^{+16.7}_{-16.3}$  & 244.3$^{+15.3}_{-15.3}$ \\
$L_{\rm IR}$\tnd              & 88.5$^{+2.6}_{-2.6}$     & 89.2$^{+2.5}_{2.5}$     \\
$M_{\rm dust}$\tne            & 50.5$^{+20.4}_{-20.2}$   & 25.7$^{+3.9}_{-5.5}$
\enddata
% if each PDF is Gaussian, mean ~ best
% else mean != best
\label{tab:mbb}
\tablenotetext{a}{Rest-frame wavelength where $\tau_\nu$\,=\,1 in $\micron$}% at which the dust becomes optically thick; upper limit is 2000 $\micron$}
\tablenotetext{b}{Observed-frame wavelength of the SED peak in $\micron$}
\tablenotetext{c}{Observed-frame flux density at 500 $\micron$ in mJy}
\tablenotetext{d}{Rest-frame 8-1000 $\micron$ luminosity in 10$^{12}$ \Lsun}
\tablenotetext{e}{Dust mass in 10$^8$ M$_\odot$, derived assuming a standard absorption mass coefficient $\kappa$=2.64 m$^2$ kg$^{-1}$ at $\lambda$=125.0 $\micron$ \citep{Dunne03a}}
\tablecomments{Errors reported here are $\pm$1$\sigma$. $L_{\rm IR}$ and $M_{\rm d}$ are reported prior to lensing correction.}
\end{deluxetable}